Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
fixed broken merge

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{amsthm}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{group-separator={,},group-minimum-digits={3},output-decimal-marker={.}}

@@

\begin{figure}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\textit{region} \equiv \textit{region}'$} \\[1em]
\begin{array}{rll}
(\textsc{F\, \texttt{+} \,F}) \;\; &
\stenFwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenFwdSR{m}{d}{s} \\[1em]
%
%
(\textsc{B\, \texttt{+} \,B}) \;\; &
\stenBwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenBwdS{n}{d}{r} \; \texttt{+} \; \stenBwdSR{m}{d}{s} \\[1em]
%
%
(\textsc{C\, \texttt{+} \,C}) \;\; &
\stenCenSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
\equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \; \stenCenSR{m}{d}{s} \\[1em]
%
%
(\textsc{C\, \texttt{+} \,F}) \;\; & \stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenFwdSR{m}{d}{s} \\[1em]
%
%
(\textsc{C\, \texttt{+} \,B}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenBwdSR{m}{d}{s} \\[1em]
%
%
(\textsc{B\, \texttt{+} \,F}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenBwdSR{n}{d}{s}
  \\[1em]
%
%
(\textsc{R\, \texttt{+} \,F}) \;\; &
\stenFwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenFwdSR{n}{d}{r} \\[1em]
%
%
(\textsc{R\, \texttt{+} \,B}) \;\; &
\stenBwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenBwdSR{n}{d}{r} \\[1em]
%
%
(\textsc{R\, \texttt{+} \,C}) \;\; &
\stenCenS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenCenSR{n}{d}{r} \\[1em]
%
%
(\textsc{irrefl}) \;\; & (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \;
\equiv \; R \, \term{*} \, S \\[0.5em]
%
%
(\texttt{+}\textsc{idem}) \;\; & S \; \texttt{+} \; S \; \equiv \; S \\[0.5em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.5em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.5em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.5em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.5em]
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)
\end{array}
%
%\framebox{${\textit{specDec} :: \texttt{v}} \equiv
%{\textit{specDec'} :: \texttt{v}}$} \\[1em]
%\begin{array}{rl}
% MUTUAL
%(\textsc{mutual}) \; &
%\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
%  :: \texttt{w}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
%  \texttt{v} \\[0.5em]
%(\textsc{coalesce}) \; &
%\texttt{stencil} \; \texttt{dependency} (\bar{v})
%  :: \texttt{v} \\
%& ; \texttt{stencil} \; \texttt{dependency} (\bar{w})
%  :: \texttt{v}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
%  \texttt{v} \\[0.5em]
%\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications}
\label{fig:equations}
\end{figure}

@@

We define a semantic model for our stencil specification language.
This model has a number of purposes: (1) it is used in the inference,
checking, and program synthesis algorithms (introduced in
Section~\ref{sec:analysis}); (2) gives a basis for the equational
theory and approximations of the previous section; and (3) can be used
to guide correct implementations. The model is a denotational
semantics, in terms of sets of vectors which we call \emph{index
  schemes}. We introduce first the notion of indexing schemes
and \emph{unification} which gives the domain of our model
and forms the basis of the rest of the algorithsm for our
specificaiton. We then define the denotational model.

\subsection{Indexing schemes and unification}

@@

\begin{definition}[Unification]
Two schemes $u, v$ of dimensionality $n$
are said to be \emph{unifiable}, written
$u \sim_n v$, iff the following holds:
%%
\[
u \sim_n v \; \Leftrightarrow  \;
  \forall i\!\in\!\{ 1,.., n \} . \, (u_i = \infty
\, \vee \, v_i = \infty \, \vee \, u_i = v_i)
\]
%%
That is, two schemes are related by the binary relation
$\sim_n$ over $(\mathbb{Z}_{\infty})^n$ if in each dimension
either is $\infty$ or they are equal. 
Thus, the element $\infty$ represents the ``general acceptor''
 meaning it unifies with any index in the same dimension.

The subscript on $\sim_n$ is sometimes dropped when the
intended dimensionality is clear or inconsequential.
\end{definition}

@@

\begin{example}
Take three array indexing terms
$(i, j+1)$, $(1, j+1)$ and $(3, j)$.
As given by \textsf{schematic}, these have the 
corresponding schemes: $\vtwoh{0}{1}$, $\vtwoh{\infty}{1}$
and $\vtwoh{\infty}{0}$ respectively. The following
relationship of unification (or lack of) then holds between
them:
%%
\begin{equation*}
\vtwoh{\infty}{0} \; \not\sim_2 \; \vtwoh{0}{1} \; \sim_2 \;
\vtwoh{\infty}{1} \; \not\sim_2 \; \vtwoh{\infty}{0}
\end{equation*}
%%
The middle pair unifies as $0$ and $\infty$ unify, 
whereas the left and right pairs do not unify as
$0 \not\sim_1 1$. 
\end{example}

@@

Let $M$ range over the finite maps generated by \textsf{analyse}.  For
each $v \in \mathsf{dom}(M)$, the algorithm then constructs a group
of regions (called \emph{spans}) which cover all contiguous groups of relative indices
(from $M(v)$) in the $n$-dimensional space.
Informally, indexing schemes are converted to pairs $u \mapsto (u,
u)$, defining the lower and upper bound of a unit span 
in $n$-dimensions. For
example, for the five point stencil there are five $1 \times 1$ spans.
%
\begin{center}
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture}
\end{center}
%
Each dimension is traversed, coalescing consecutive spans
which vary in $m$-dimensions into coalesced spans which vary in
$m+1$-dimensions. In this example, the $0$-dimensional spans
 become $1$-dimensional spans (rows/columns):
%
\begin{align*}
\begin{array}{cc}
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\end{tikzpicture}
\qquad
&
\qquad
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1.6,1) rectangle (2.1,1.5);
\draw[thick] (0.4,1) rectangle (0.9,1.5);
\draw[thick] (1,0.5) rectangle (1.5,2);
\end{tikzpicture}
\\
\textit{dimension 1} \qquad & \qquad \textit{dimension 2}
\end{array}
\end{align*}
%
Any spans that are contained with any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another).
\begin{center}
\begin{tikzpicture}
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (2,1.22) {+};
\draw[thick] (2.5,1) rectangle (4,1.5);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture}
\end{center}
%
This procedure is iterated till a fixed point is reached. In this
example this is reached in the first step. The final spans
are used to construct stencil specification, where
each span is product \texttt{*} of regions and multiple spans are
combined with the region sum \texttt{+}.

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{group-separator={,},group-minimum-digits={3},output-decimal-marker={.}}
\usepackage{tikz}
\usepackage{xypic}

\tikzset{round left paren/.style={ncbar=0.4cm,out=110,in=-110}}
\tikzset{round right paren/.style={ncbar=0.4cm,out=70,in=-70}}

@@

\begin{figure}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\textit{region} \equiv \textit{region}'$} \\[0.7em]
\begin{array}{rll}
(\textsc{F\, \texttt{+} \,F}) \;\; &
\stenFwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,B}) \;\; &
\stenBwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenBwdS{n}{d}{r} \; \texttt{+} \; \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,C}) \;\; &
\stenCenSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
\equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \; \stenCenSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,F}) \;\; & \stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,B}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,F}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenBwdSR{n}{d}{s}
  \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,F}) \;\; &
\stenFwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenFwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,B}) \;\; &
\stenBwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenBwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,C}) \;\; &
\stenCenS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenCenSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{irrefl}) \;\; & (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \;
\equiv \; R \, \term{*} \, S \\[0.5em]
%
%
(\texttt{+}\textsc{idem}) \;\; & S \; \texttt{+} \; S \; \equiv \; S \\[0.5em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.5em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.5em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.5em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.5em]
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)
\end{array}
%
%\framebox{${\textit{specDec} :: \texttt{v}} \equiv
%{\textit{specDec'} :: \texttt{v}}$} \\[0.7em]
%\begin{array}{rl}
% MUTUAL
%(\textsc{mutual}) \; &
%\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
%  :: \texttt{w}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
%  \texttt{v} \\[0.5em]
%(\textsc{coalesce}) \; &
%\texttt{stencil} \; \texttt{dependency} (\bar{v})
%  :: \texttt{v} \\
%& ; \texttt{stencil} \; \texttt{dependency} (\bar{w})
%  :: \texttt{v}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
%  \texttt{v} \\[0.5em]
%\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications}
\label{fig:equations}
\end{figure}

@@

We define a semantic model for our stencil specification language.
This model has a number of purposes: (1) it is used in the inference,
checking, and program synthesis algorithms (introduced in
Section~\ref{sec:analysis}); (2) gives a basis for the equational
theory and approximations of the previous section; and (3) can be used
to guide correct implementations. The model is a denotational
semantics, in terms of sets of vectors which we call \emph{index
  schemes}. We introduces these first then define the denotational semantics.

\subsection{Indexing schemes and generalisation}

@@

\begin{definition}[Generalisation]
Given two schemes $v, w$ of dimensionality $n$,
the $w$ is said to \emph{generalise} $v$ 
(or conversely $v$ \emph{specialises} $w$),
written $v \preceq_n w$, iff the following holds:
%%
\[
v \preceq_n w \;\; \Leftrightarrow  \;\;
  \forall i\!\in\!\{ 1, \ldots, n \} . \, (w_i = \infty \, \vee \, v_i = w_i)
\]
%%
That is, two schemes are related by the (reflexive and transitive) binary relation
$\preceq_n$ over $(\mathbb{Z}_{\infty})^n$ if in each dimension
the two schemes are equal or $w_i = \infty$ 
in place of some $\mathbb{Z}$ in $v_i$ for $v \preceq_n w$. 
Thus, the element $\infty$ 
generalising any index in the same dimension.

The subscript on $\preceq_n$ is sometimes dropped when the
intended dimensionality is clear or inconsequential.
\end{definition}

@@

\begin{example}
Take three array indexing terms
$(i, j+1)$, $(1, j+1)$ and $(3, j)$.
As given by \textsf{schematic}, these have the 
corresponding schemes: $\vtwoh{0}{1}$, $\vtwoh{\infty}{1}$
and $\vtwoh{\infty}{0}$ respectively. The following
relationship of generalises then holds:
%\[
%\xymatrix@R=0.2em@C=2em{
%[\infty \;\; 0] \ar@{-}[dr]^{\preceq_n} &  \\
%& [0 \;\; 1]  \\
%[\infty \;\; 1]  \ar@{-}[ur]_{\preceq_n} & 
%}
%\]
%%
\begin{equation*}
\vtwoh{\infty}{0} \; \not\preceq_2 \; \vtwoh{0}{1} 
\qquad
\vtwoh{\infty}{1} \; \preceq_2 \; \vtwoh{0}{1} 
\end{equation*}
%%
\end{example}

@@

Let $M$ range over the finite maps generated by \textsf{analyse}.  For
each $v \in \mathsf{dom}(M)$, the algorithm then constructs a group
of regions (called \emph{spans}) which cover all contiguous groups of relative indices
(from $M(v)$) in the $n$-dimensional space.
Informally, indexing schemes are converted to pairs $u \mapsto (u,
u)$, defining the lower and upper bound of a unit span 
in $n$-dimensions. For
example, for the five point stencil there are five $1 \times 1$ spans.
Figure~\ref{fig:inference-steps-informal}(a) gives an illustration
of these five unit spans.

\begin{figure}[t]
\vspace{-0.3em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0.2em]
{\scalebox{0.75}{
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} 
}} \\[0.36em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
{\scalebox{0.75}{
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture}
}}
 \\ (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
{\scalebox{0.75}{
\begin{tikzpicture}
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture}
}}  \\ (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in 
specification inference}
\label{fig:inference-steps-informal}
\end{figure}

%
Each dimension is traversed, coalescing consecutive spans
which vary in $m$-dimensions into coalesced spans which vary in
$m+1$-dimensions. In this example, the $0$-dimensional spans
 become $1$-dimensional spans (rows/columns), illustrated
in Figure~\ref{fig:inference-steps-informal}(b).
Any spans that are contained with any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another) (Figure~\ref{fig:inference-steps-informal}(c).

This procedure is iterated till a fixed point is reached. In this
example this is reached in the first step. The final spans
are used to construct stencil specification, where
each span is product \texttt{*} of regions and multiple spans are
combined with the region sum \texttt{+}.
